Extended Formal Analysis of the EDHOC Protocol in Tamarin

### 2022-07-22 intro.tex

%-------------------------------------------------------------------------- sub
\subsection{Contributions}
\label{sec:contributions}
In this paper, we extend the work presented in~\cite{Norr21}. 
%
We analyze the version of \mEdhoc{} as in~\cite{}.
%
We extend the adversary model and \mTamarin{} models to capture weak
post-compromise security (PCS) and model Trusted Execution Environments
(TEE).
%
We also modify the model to consider a stronger version of the \mXor{} encryption whenever possible. 
%
We check for the following properties: 
\begin{itemize}
\item Perfect forward secrecy for the session key material
\item Injective agreement
\item Implicit agreement
\item Weak post-compromise security
\item Secrecy and integrity of \mADthree
\end{itemize}
\vnote{Check this!}
%
We also discuss various issues arising in relation to the use of trusted 
execution environments (TEEs), denial of service (DoS) attacks, sources of 
randomness, the negotiation of parameters (which the formal model abstracts 
away) for establishing the protocol, and other potential attacks and concerns.
%
We have communicated these issues to the developers of the protocol.

>>>>

%-------------------------------------------------------------------------- sub
\subsection{Contributions}
\label{sec:contributions}
In this paper, we extend the work presented in~\cite{Norr21}.
%
We analyze the version of \mEdhoc{} as in~\cite{}.
%
We extend the adversary model and \mTamarin{} system models to capture weak
post-compromise security (PCS) and model Trusted Execution Environments
(TEE).
%
We check for the following properties:
\begin{itemize}
\item Injective agreement
\item Implicit agreement
\item Weak post-compromise security for the session key material
\item Perfect Forward Secrecy (PFS) for the session key material
\item Secrecy and integrity of \mADthree
\end{itemize}
%
We follow the definition of weak PCS by~\cite{DBLP:conf/csfw/Cohn-GordonCG16},
which subsumes PFS.
%
We also discuss various issues arising in relation to the use of trusted
execution environments (TEEs), denial of service (DoS) attacks, error handling,
the negotiation of parameters (which the formal model abstracts
away) for establishing the protocol, and other potential attacks and concerns.
%
We have communicated these issues to the developers of the protocol.
%

### 2022-07-22 kdfdiagram.tex

\draw [->, norm] (p3.south) ++(0,-0.5) -- (h8);
\draw [->, norm] (h8) -- (k2b);
\draw [->, norm] (p2) -- (h2); 
\draw [->, norm] (c1.east) -- ++(1.85,0) -- (h3.north) node[midway,above left] {Y};
%\draw [->, norm] (h3.south) -- ++(0,-1) -- ++(-3,0);
\draw [->, norm] (h3.west) -- (p3.east);
\draw [->, norm] (p3) -- (h4); 
\draw [->, norm] (c2.east) -- ++(1.91,0) -- (h5.north) node[midway,above left] {Y};
%\draw [->, norm] (h5.south) -- ++(0,-1) -- ++(-3,0);
\draw [->, norm] (h5.west) -- (p4.east);
\draw [->, norm] (p4) -- (h6);
\draw [->, norm] (p4.west) ++(-0.25,-0) -- ++(0,-0.8) -- (h9.east);

\node [hkdfext, right=3cm of p2] (h1) {\mHkdfExtract};
\node [dhk, right=2.7cm of h1] (p0) {$\mGxy$};
\node [terminput, text width=2em, below = 0.2cm of p0] (u1) {Salt};
\draw [->] (h1.west) -- (p2.east);
\draw [->] (u1.west) -- ++(-1.52,0) -- (h1.south);
\draw [->] (p0.west) -- (h1.east);

>>>>

\draw [->, norm] (p3.south) ++(0,-0.5) -- (h8);
\draw [->, norm] (h8) -- (k2b);
\draw [->, norm] (p2) -- (h2); 
\draw [->, norm] (c1.east) -- ++(2.025,0) -- (h3.north) node[midway,above left] {Y};
%\draw [->, norm] (h3.south) -- ++(0,-1) -- ++(-3,0);
\draw [->, norm] (h3.west) -- (p3.east);
\draw [->, norm] (p3) -- (h4); 
\draw [->, norm] (c2.east) -- ++(2.07,0) -- (h5.north) node[midway,above left] {Y};
%\draw [->, norm] (h5.south) -- ++(0,-1) -- ++(-3,0);
\draw [->, norm] (h5.west) -- (p4.east);
\draw [->, norm] (p4) -- (h6);
\draw [->, norm] (p4.west) ++(-0.25,-0) -- ++(0,-0.8) -- (h9.east);

\node [hkdfext, right=3cm of p2] (h1) {\mHkdfExtract};
\node [dhk, right=2.7cm of h1] (p0) {$\mGxy$};
\node [terminput, text width=2em, below = 0.2cm of p0] (u1) {Salt};
\draw [->] (h1.west) -- (p2.east);
\draw [->] (u1.west) -- ++(-2.27,0) -- (h1.south);
\draw [->] (p0.west) -- (h1.east);

### 2022-07-22 paper.tex

%----------------------------------------------------------------------- subsub
\subsubsection{Denial of Service.}
Denial of Service (DoS) attacks for IoT protocols are well-studied,
but DoS aspects particular to \mEdhoc{} is not.
%
The \mSpec{} does not include any specific measures for countering such 
attacks,
but assumes the application takes care of it.

@@

%------------------------------------------------------------------------- sub
\subsection{Unclear Security Model}
In addition to unclear contexts in which
\mEdhoc{} can be used, the \mSpec{} also does not provide enough detail 
about
adversary capabilities.
%
The four different cryptographic cores included in \mEdhoc{} are based on
academic protocols which are designed to work with highly specific but
potentially different adversary models.
%
%For example, \mOptls{}, which essentially has the same core as \mStat{}, is
%shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, 
%which
%explicitly separates long-term key storage from that of the ephemeral keys 
%and
%session state.
%%
%This fits in well with our use of TEEs.
Since \mSigma{} cannot protect against compromised ephemeral keys, the 
authors
of \mEdhoc{} felt that considering the compromise of ephemeral keys 
separately
from that of long-term keys was not required~\cite{personalCommunication}.
%
The reason was presumably based on the fact that the \mSigSig{} method was
modeled closely on the \mSigmaI{} variant of \mSigma{}, and that it would be
preferable to obtain a homogeneous security level among the \mEdhoc{}
methods.~\cite{Norr21}
%
However, this only holds true for situations where one is only interested in 
the
session key confidentiality for an ongoing session.
%
Secure modules, in addition to being able to store long-term keys, are useful 
in
other ways too.
%
They also provide weak Post-Compromise Security (PCS) guarantees as 
discussed
above.
%
Upon discussion with the authors, they included recommendations on how to 
store
long-term keys, and how to perform operations on these keys inside a secure
module.
%

@@

\subsection{A Note About the Version of \mEdhoc}
\label{sec:newdrafts}
In this work, we have analyzed the \mEdhoc{} specification as of July
2020~\cite{our-analysis-selander-lake-edhoc-00}.
%
While our formal analysis applies to at least the version from February 2021, there are newer versions.
%
The latest version is from July 2022~\cite{draft-ietf-lake-edhoc-15}, which, among other things, changes the key derivation function.
%
However, we do refer to various aspects (error handling, denial of service etc) of this latest version, in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.

%

\knote{we 
%are already citing too many different versions, so I don't think we
    should add this to the bib. Potentially, we should tomorrow instead remove
    and just mention some of the later versions.}
\vnote{We do need to cite the latest one, just so people can look it up. It would 
be more sensible to remove the citation for the one from February 2021, and 
instead be more concrete about which version of EDHOC is the latest one that 
our formal analysis applies to.}
%

>>>>

%----------------------------------------------------------------------- subsub
\subsubsection{Denial of Service.}
Denial of Service (DoS) attacks for IoT protocols are well-studied,
but DoS aspects particular to \mEdhoc{} is not.
%
The \mSpec{} does not include any specific measures countering such
attacks, but assumes the application takes care of it.

@@

%------------------------------------------------------------------------- sub
\subsection{Unclear Security Model}
In addition to unclear contexts in which
\mEdhoc{} can be used, the \mSpec{} also does not provide enough detail about
adversary capabilities.
%
The four different cryptographic cores included in \mEdhoc{} are based on
academic protocols which are designed to work with highly specific but
potentially different adversary models.
%
%For example, \mOptls{}, which essentially has the same core as \mStat{}, is
%shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, which
%explicitly separates long-term key storage from that of the ephemeral keys and
%session state.
%%
%This fits in well with our use of TEEs.
Since \mSigma{} cannot protect against compromised ephemeral keys, the authors
of \mEdhoc{} felt that considering the compromise of ephemeral keys separately
from that of long-term keys was not required~\cite{personalCommunication}.
%
The reason was presumably based on the fact that the \mSigSig{} method was
modeled closely on the \mSigmaI{} variant of \mSigma{}, and that it would be
preferable to obtain a homogeneous security level among the \mEdhoc{}
methods.~\cite{Norr21}
%
However, this only holds true for situations where one is only interested in the
session key confidentiality for an ongoing session.
%
Secure modules, in addition to being able to store long-term keys, are useful in
other ways too.
%
They also provide weak Post-Compromise Security (PCS) guarantees as discussed
above.
%
Upon discussion with the authors, they included recommendations on how to store
long-term keys, and how to perform operations on these keys inside a secure
module.
%

@@

\subsection{A Note About the Version of \mEdhoc}
\label{sec:newdrafts}
In this work, we have analyzed the \mEdhoc{} specification as of July
2020~\cite{our-analysis-selander-lake-edhoc-00}.
%
While our formal analysis applies to v5 of the \mSpec{} (February 2021),
there are newer versions.
%
The latest version at the time of writing is
v15 (July 2022)~\cite{draft-ietf-lake-edhoc-15}, which, among other things,
differ from v5 in terms of some inputs to the key derivations.
%
However, we do refer to various aspects
(error handling, denial of service etc) of this latest version,
in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.
%
Due to lack of time we unfortunately could not bring the Tamarin model up to
date with v15~\cite{draft-ietf-lake-edhoc-15}.
%
